\large{\bf Looking forwards and backwards through locally regulated populations }

### 2023-04-17 forwards-and-backwards.tex
updated figs

\begin{figure}
    \begin{center}
        \includegraphics[width=0.45\textwidth]{figures/ex2a/pme_1234.steps}
        \includegraphics[width=0.45\textwidth]{figures/ex2b/pme_1234.steps}
    \end{center}
    \caption{
        Two panels, showing simulated 1D populations under a porous medium equation with logistic growth~(\ref{PME}),
        % comparing the simulated profile to the analytical solution;
        $\theta/N$ small on the left; large on the right.
        Simulation is with $\gamma(m) = m$ and $\mu(m) = (1+1/\theta) m - 1/\theta$
        (implemented by choosing a small $dt$ taking reproduction probability per time step
        to be $1 - \exp(- dt m)$
        and death probability to be $\mu(m) = 1 - \exp(-dt ((1+1/\theta) m - 1/\theta))$).
        \label{fig:pme_waves}
    }
\end{figure}

In one dimension, equation~(\ref{PME}) has an explicit travelling wave solution
\begin{align} \label{eqn:pme_wave}
    w^P(t, x)
    :=
    \left( 1 - e^{ \frac{1}{2} (x - x_0 - t) } \right)_+ .
\end{align}
Notice that the wave profile has a sharp boundary at $x = x_0 + t$.
There are also travelling wave solutions with $c>1$ \citep{gilding/kersner:2005},
which lack this property.
However, for initial conditions that decay sufficiently rapidly at infinity,
such as one might use in modelling a population invading new territory,
the solution converges to \eqref{eqn:pme_wave} \citep{kamin/rosenau:2004}.
In Figure~\ref{fig:pme_waves} we show simulations of the individual based model
described above. They
display travelling wave solutions
similar to numerical solutions of~(\ref{PME}),
with increasingly good agreement for larger $\theta$ and $N$.
\comment{TODO: does it?}
As expected, noise slows down the wave.
\comment{Not sure that noise will slow things drastically because the 
biggest production of
mass happens behind the wavefront.}

>>>>

\begin{figure}
    \begin{center}
        \includegraphics[width=0.45\textwidth]{figures/ex2a/pme_1234.steps}
        \includegraphics[width=0.45\textwidth]{figures/ex2b/pme_1234.steps}
    \end{center}
    \caption{
        Two panels, showing simulated 1D populations under a porous medium equation with logistic growth~(\ref{PME}) in $d=1$,
        % comparing the simulated profile to the analytical solution;
        $\theta$ small on the left; large on the right.
        Simulation is with $\gamma(m) = m$ and $\mu(m) = (1+1/\theta) m - 1/\theta$
        (implemented by choosing a small $dt$ taking reproduction probability per time step
        to be $1 - \exp(- dt m)$
        and death probability to be $\mu(m) = 1 - \exp(-dt ((1+1/\theta) m - 1/\theta))$);
        values of $\theta$ are 1 and 100, respectively.
        \label{fig:pme_waves}
    }
\end{figure}

In one dimension, equation~(\ref{PME}) has an explicit travelling wave solution
\begin{align} \label{eqn:pme_wave}
    w^P(t, x)
    :=
    \left( 1 - e^{ \frac{1}{2} (x - x_0 - t) } \right)_+ .
\end{align}
Notice that the wave profile has a sharp boundary at $x = x_0 + t$.
There are also travelling wave solutions with $c>1$ \citep{gilding/kersner:2005},
which lack this property.
However, for initial conditions that decay sufficiently rapidly at infinity,
such as one might use in modelling a population invading new territory,
the solution converges to \eqref{eqn:pme_wave} \citep{kamin/rosenau:2004}.
In Figure~\ref{fig:pme_waves} we show simulations of the individual based model described above.
They display travelling wave solutions qualitatively similar to solutions of~(\ref{PME}),
with increasingly good agreement for larger $\theta$ and $N$.
\comment{TODO: does it?}
As expected, noise slows down the wave.
\comment{Not sure that noise will slow things drastically because the 
biggest production of
mass happens behind the wavefront.}
